---
title: "  **Craiglist-Used_Cars-Analysis**"
output: pdf_document
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{setspace}\spacing{1.5}
  - \usepackage{titling}
  - \posttitle{\end{center}}
  - \usepackage{sectsty} \sectionfont{\emph}
---
\begin{centering}
\vspace{1cm}
\large
{\bf Submitted By : David Joseph Johnson } \\
\vspace{.75cm}
{\bf 02/20/2022 } \\
\end{centering}
\newpage
\tableofcontents
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Craigslist is an American classified advertisement website with sections for jobs, gigs, housing, vehicles etc. for sale. The dataset used here is scraped from data collected specifically from the used vehicle's sections of Craigslist.

## SOURCE
The Dataset was retrieved from Kaggle (online community for Data Science Enthusiasts).

## DASHBOARD
The aim of the dashboard is to provide an interactive aid to compare the vehicles listed on Craigslist.
The dashboard consists of 2 different visualizations interactive visualizations:
• Vehicle Count by Manufacturer: Shows popular brands in the used vehicle sales
• Price Variation Based on Type: Shows average resale price based on the type of the car.
Filters Transmission and Fuel can be toggled to view respective results



# Code

## Invoking Required Libraries
```{r , echo=FALSE ,results='hide', warning=FALSE, message=FALSE}
### Install ISLR Package which contains the College Dataset
### Invoke Packages
library(shiny)
library(shinydashboard)
library(DataExplorer)
library(scales)
library(dplyr)
library(ggplot2)
library(corrplot)
library(dplyr)
library(janitor)
library(tidyr)
library(RColorBrewer)
library(car)
library(psych)
library(imputeTS)
library(corrplot)
```

## Importing Dataset
```{r }
raw_df <- read.csv("vehicles.csv",header = TRUE)
```

## Data Cleaning
```{r , echo=FALSE ,results='hide', warning=FALSE, message=FALSE}
df<- raw_df
df<-clean_names(df)
df<- subset(df,select=- c(url,cylinders,drive,description,size,vin,lat,long,region_url,image_url))
df<-df %>% remove_empty(which = c("rows"))
df<-df %>% remove_empty(which = c("cols"))
df<-na.omit(df)
df<- df[!(is.na(df$paint_color) | df$paint_color==""),]
df<- df[!(is.na(df$type) | df$type==""),]
df<- df[!(is.na(df$id) | df$id==""),]
df<- df[!(is.na(df$region) | df$region==""),]
df<- df[!(is.na(df$price) | df$price==""),]
df<- df[!(is.na(df$year) | df$year==""),]
df<- df[!(is.na(df$manufacturer) | df$manufacturer==""),]
df<- df[!(is.na(df$model) | df$model==""),]
df<- df[!(is.na(df$condition) | df$condition==""),]
df<- df[!(is.na(df$fuel) | df$fuel==""),]
df<- df[!(is.na(df$odometer) | df$odometer==""),]
df<- df[!(is.na(df$title_status) | df$title_status==""),]
df<- df[!(is.na(df$transmission) | df$transmission==""),]
df<- df[!(is.na(df$type) | df$type==""),]
df<- df[!(is.na(df$paint_color) | df$paint_color==""),]
df<- df[!(is.na(df$state) | df$state==""),]
df<- df[!(is.na(df$posting_date) | df$posting_date==""),]
df<- df[order(df$posting_date),]
df<- df %>% top_n(-10000)
```

\newpage

## RShiny Dashboard

```{r }
# Uncomment next line to import Cleaned Dataset directly to skip data cleaning process and start dashboard using this single chunk of code
 df <- read.csv("cleaned_vehicles.csv",header = TRUE)
#Set Dashboard Title
header <- dashboardHeader(title = "Quote Right", titleWidth=150)  

#Sidebar content of the dashboard
sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Overview", tabName = "dashboard", icon = icon("glyphicon glyphicon-list",lib='glyphicon')
             )
    )
  )
  
flash_cards <- fluidRow(valueBoxOutput("card1"),valueBoxOutput("card2"),valueBoxOutput("card3"))


graphs <- fluidRow(
  box(
    title = "Vehicle Count By Manufacturer"
    ,status = "primary"
    ,solidHeader = TRUE 
    ,collapsible = TRUE 
    ,plotOutput("manufacturer_count", height = "700px", width = "580px")
  ),
  
  #Filters using radio buttons
  box(
    title = "Filters"
    ,status = "primary"
    ,solidHeader = TRUE 
    ,collapsible = TRUE 
    ,radioButtons("x", "TRANSMISSION",
                 list("All"='a', "Automatic"='b', "Manual"='c'), inline=T),
    
    radioButtons("y", "FUEL",
                 list("All"='d', "Gas"='e', "Diesel"='f'), inline=T), height = "170px"
    ),

  
  box(
    title = "Average Price By Type"
    ,status = "primary"
    ,solidHeader = TRUE 
    ,collapsible = TRUE 
    ,plotOutput("vehicle_type", height = "510px",width = "580px")
  )
)


# Combining Fluid rows into the dashboard body
body <- dashboardBody(flash_cards, graphs)

# Dashboard Page UI
ui <- dashboardPage(title = 'Quote Right Dashboard', header, sidebar, body, skin='blue')

# Creating server to facilitate functions in UI 
server <- shinyServer(function(input, output) { 
  

  # Extracting Unique Manufacturers
  total.manufacturers <- length(unique(df$manufacturer))
  # Flashcard1: Total Unique Manufacturers
  output$card1 <- renderValueBox({
    valueBox(
      formatC(total.manufacturers, format="d", big.mark=',')
      ,'Total Number of Manufacturers'
      ,icon = icon("fa-solid fa-car")
      ,color = "light-blue")
  })
  
  # Computing Average Price
  average.pricing <- mean(df$price)
  # Flashcard2: Average Price
  output$card2 <- renderValueBox({
    valueBox(
      formatC(average.pricing, format="d", big.mark=',')
      ,'Average Price'
      ,icon = icon("glyphicon glyphicon-usd",lib='glyphicon')
      ,color = "light-blue")
  })
  
  # Computing Average Odometer
  average.odometer <- mean(df$odometer)
  # Flashcard3: Average Odometer
  output$card3 <- renderValueBox({
    valueBox(
      formatC(average.odometer, format="d", big.mark=',')
      ,'Average Odometer'
      ,icon = icon("glyphicon glyphicon-road",lib='glyphicon')
      ,color = "light-blue")
  })
  

    filtered_df <- df
  
# Barplot of Count of Vehicles by Manufacturers
  output$manufacturer_count <- renderPlot({
    filtered_df <- df
     if(req(input$x)=='b'){
      filtered_df <- filter(df, transmission == "automatic")
    }
    if(req(input$x)=='c'){
      filtered_df <- filter(df, transmission == "manual")
    }
    if(req(input$y)=='e'){
     filtered_df <- filter(filtered_df, fuel == "gas")
    }
    if(req(input$y)=='f'){
      filtered_df <- filter(filtered_df, fuel == "diesel")
    }
  ggplot(filtered_df, aes(x = forcats::fct_rev(forcats::fct_infreq(manufacturer)))) +
  theme(axis.text.x = element_text(face="bold", color="black", 
                           size=10),
          axis.text.y = element_text(face="bold", color="black", 
                           size=10)) + 
  theme(plot.caption = element_text(hjust = 0.1)) + 
  geom_text(aes(label = ..count..), stat = "count", hjust = -0.1,size=4, colour = "black") + 
  geom_bar(fill="lightblue")  + ylim(0, 1100) +  ylab("Count") + 
      xlab("Manufacturer")+
  coord_flip()
})
  
  # Barplot of Average Price By Type
  output$vehicle_type <- renderPlot({
    filtered_df <- df
    if(req(input$x)=='b'){
      filtered_df <- filter(df, transmission == "automatic")
    }
    if(req(input$x)=='c'){
      filtered_df <- filter(df, transmission == "manual")
    }
    if(req(input$y)=='e'){
     filtered_df <- filter(filtered_df, fuel == "gas")
    }
    if(req(input$y)=='f'){
      filtered_df <- filter(filtered_df, fuel == "diesel")
    }
    
  ggplot(filtered_df, aes(x = forcats::fct_rev(forcats::fct_infreq(type))),y= mean(price)) +
  theme(axis.text.x = element_text(face="bold", color="black", 
                           size=10,angle = 45),
          axis.text.y = element_text(face="bold", color="black", 
                           size=10)) + 
  theme(plot.caption = element_text(hjust = 0.1)) + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -1,size=4, colour = "black") + 
  geom_bar(fill="lightblue")  + ylim(0, 2700) +  ylab("Average Price ($)") + 
      xlab("Type")
})
}
)
shinyApp(ui, server)
```



# References

* https://cran.r-project.org/web/packages/shinydashboard/shinydashboard.pdf

* ggplot2, R., DuBois, C., Chang, J., & Pauloo, R. (2009). Rotating and spacing axis labels in ggplot2. Retrieved 20 February 2022, from https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2

* Shiny?, H. (2018). How to extract the values of dynamically generated inputs in Shiny?. Retrieved 20 February 2022, from https://stackoverflow.com/questions/50795355/how-to-extract-the-values-of-dynamically-generated-inputs-in-shiny